GPU-friendly Stroke Expansion

### 2024-04-16 paper.tex
Apply sigconf, authordraft template
Address feedback

\documentclass[format=acmsmall]{acmart}
\usepackage[linesnumbered]{algorithm2e}

\begin{document}
\title{GPU-friendly Stroke Expansion}
\author{Raph Levien}
\author{Arman Uguray}
\affiliation{
    \institution{Google}
    \city{San Francisco}
    \state{CA}
    \country{USA}
}

@@

We aimed to satisfy the following the criteria:
\begin{enumerate}
  \item The implementation must be able to handle a large number of inputs at real-time rates. 
  \item The stroke outlines must meet the weak correctness definition.
  \item The implementation must support the standard SVG stroke end cap (\emph{butt}, \emph{square}, \emph{round}) and join styles (\emph{bevel}, \emph{miter}, \emph{round}).
  \item The CPU must do no work beyond the basic preperation of the input path data for GPU consumption.
\end{enumerate}

@@

Our compute shader is arranged as a 1-dimensional grid, parallelized over input path segments. We chose a workgroup size of 256 as it works well over a wide range of GPUs. Each thread in the global grid is assigned to process a single path segment. We encode the individual path segments contiguously in a GPU storage buffer such that each element has 1:1 correspondance to a global thread ID in the dispatch. This layout easily scales to tens of thousands of input path segments, which can be distributed across any number of paths and subpaths.

% TODO: caps/joins figure highlighting mapping of segments, joins, and caps to threads, encoding.
% TODO: visualize per-segment flattened output for a fill and for a stroke from the perspective of one thread.
% TODO: finish this \begin{algorithm} snippet
\begin{algorithm}
  $tid \gets$ global invocation ID\;
  $(style, segment) \gets readScene(tid)$\;
  \eIf{$style$ is $fill$}{
      flatten source curve
  }{\If{you're sad}{
      just keep going
  }}
\caption{The control flow of our compute shader}
\end{algorithm}

Each thread outputs a polyline that fits a subset of the rendered subpath's outer contour, represented by the path segment. For a filled primitive, a thread only outputs the flattened approximation of the path segment. For a stroke, a thread outputs the polylines approximating multiple curves: a) the two parallel curves on both sides of the source segment, offset by the desired stroke \emph{halfwidth}, and b) the inner and outer joins that connect the path segment to the next adjacent segment. We also store metadata in our input encoding that marks the position of a path segment within the containing subpath. We use this information to determine whether to output an end cap instead of a join. We encode one additional segment designated as the \emph{stroke cap marker} for every subpath. The thread assigned to the marker segment only outputs a start cap, which completes the stroke outline.

The joins between adjacent segments need to form a continuous outline when combined with the segments. This requires that a thread have access to the tangent vector at the end of its assigned segment \emph{and} the start tangent of the next adjacent segment. We require that the input segments within a subpath are stored in order which simplifies the access to the adjacent segment down to a simple read at the next array offset. Co-operation across threads is not necessary.

@@

% TODO: Talk about the inner join here. We detect inner vs outer outline based on join tangent cross product. We just draw a straight line for the inner join.
% TODO: line soup is bump allocated. Mention this here? I'm leaning towards discussing that in the Limitations section below, which will focus on memory allocation and estimation.
% TODO: while parallel on input, sequential per curve. However iterations are bounded (? needs measurement). Should this discussion happen here or down in the Performance section?

@@

% TODO: We integrated our algorithm into a fully GPU-driven rendering pipeline, Vello WGSL
% TODO: 3 alternative implementations: cubic->ES (for load balancing eval), cubic->ES->LineSoup (current), cubic->ES->ArcSoup
% TODO: GPUs tested, data sets. Test strategy
% TODO: divergence (reference prim count)

@@

% TODO: mention dashing as future work?

>>>>

\documentclass[sigconf, authordraft]{acmart}
\usepackage[linesnumbered]{algorithm2e}
\begin{document}
\title{GPU-friendly Stroke Expansion}
\author{Raph Levien}
\author{Arman Uguray}
\affiliation{
    \institution{Google}
    \city{San Francisco}
    \state{CA}
    \country{USA}
}

@@

We aimed to satisfy the following the criteria:
\begin{enumerate}
  \item The implementation must be able to handle a large number of inputs at real-time rates. 
  \item The stroke outlines must meet the weak correctness definition.
  \item The implementation must support the standard SVG stroke end cap (\emph{butt}, \emph{square}, \emph{round}) and join styles (\emph{bevel}, \emph{miter}, \emph{round}).
  \item The CPU must do no work beyond the basic preparation of the input path data for GPU consumption.
\end{enumerate}

@@

Our compute shader is arranged as a 1-dimensional grid, parallelized over input path segments. We chose a workgroup size of 256 as it works well over a wide range of GPUs. However, this choice of workgroup size is not particularly important for our algorithm as it does not require any cross-thread communication. Each thread in the global grid is assigned to process a single path segment. We encode the individual path segments contiguously in a GPU storage buffer such that each element has 1:1 correspondance to a global thread ID in the dispatch. This layout easily scales to hundreds of thousands of input path segments, which can be distributed across any number of paths and subpaths.

% TODO: caps/joins figure highlighting mapping of segments, joins, and caps to threads, encoding. Make sure to show the straight-line inner join.
% TODO: visualize per-segment flattened output for a fill and for a stroke from the perspective of one thread.
% TODO: finish this \begin{algorithm} snippet
\begin{algorithm}
  $tid \gets$ global invocation ID\;
  $(style, segment) \gets readScene(tid)$\;
  \eIf{$style$ is $fill$}{
      flatten source curve
  }{\If{you're sad}{
      just keep going
  }}
\caption{The control flow of our compute shader}
\end{algorithm}

Each thread outputs a polyline that fits a subset of the rendered subpath's outer contour, represented by the path segment. For a filled primitive, a thread only outputs the flattened approximation of the path segment. For a stroke, a thread outputs the polylines approximating multiple curves: a) the two parallel curves on both sides of the source segment, offset by the desired stroke \emph{halfwidth}, and b) the joins that connect the path segment to the next adjacent segment. We also store metadata in our input encoding that marks the position of a path segment within the containing subpath. We use this information to determine whether to output an end cap instead of a join. We encode one additional segment designated as the \emph{stroke cap marker} for every subpath. The thread assigned to the marker segment only outputs a start cap, which completes the stroke outline.

The joins between adjacent segments need to form a continuous outline when combined with the segments. This requires that a thread have access to the tangent vector at the end of its assigned segment \emph{and} the start tangent of the next adjacent segment. We require that the input segments within a subpath are stored in order which simplifies the access to the adjacent segment down to a buffer read at the next array offset. Notably, this scheme does not cooperation across threads. While we take care to output the \emph{outer} join segment (where the offset curves meet at a reflex angle) we output the inner join as a line segment.

@@

% TODO: line soup is bump allocated. Mention this here? I'm leaning towards discussing that in the Limitations section below, which will focus on memory allocation and estimation.
% TODO: while parallel on input, sequential per curve. However iterations are bounded (? needs measurement). Should this discussion happen here or down in the Performance section?
%    raph@: I don't have strong feelings on where the discussion happens. However, we want to make the following points:
%         * The concept is currently sequential only for cubic->ES, all else is parallelizable
%         * The current impl is sequential per curve, but that's just an impl choice. Personally I'm a fan of discussing this in future work related to work graphs, as I consider load balancing to be a detail
%         * Parallel lines per ES makes sense but parallel arcs per ES much less so, just because of the prim count. Thus, such detailed discussion may make sense downstream of the prim count results.

@@

% TODO: We integrated our algorithm into a fully GPU-driven rendering pipeline, Vello WGSL
% TODO: 3 alternative implementations: cubic->ES->LineSoup (current), cubic->ES->ArcSoup, less importantly: cubic->ES (for load balancing eval).
% TODO: Take a measurement with ablated bbox atomic
% TODO: discuss: GPUs tested, data sets. Test strategy
% TODO: discuss: divergence (reference prim count)

@@

% TODO: mention dashing as future work? raph@: I think so. When we describe dashing, it would be good to say that the ES representation is already optimal for dashing, as it's parametrized by arc length.
